A Monte Carlo Synthetic-Acceleration Method for Solving the Thermal Radiation Diffusion Equation\tnoteref{ornl-cr}

### 2013-02-19 doc/jcomp2012/jcomp2012.tex
adding grace plots for direct and adjoint convergence

A timing and convergence study is used to demonstrate the
effectiveness of the adjoint method as compared to the direct
method. To assess both the CPU time and number of iterations required
to converge to a solution, a problem of constant $\Delta$ was used
with varying values of grid size, fixing the spectral radius of the
system at a constant value for each variation. Both the five-point and
nine-point stencils were used with both the direct and adjoint
solvers. For each case, $N \times N$ total random walk permutations
were computed per MCSA iteration where $N \times N$ is the number of
discrete grid points in the system. Solver parameters were set to a
weight cutoff of \sn{1}{-4} for the stochastic linear solver and a
convergence tolerance of \sn{1}{-8} for the MCSA iterative solver. All
computations presented in this section were completed on a 2.66 GHz
Intel Core i7 machine with 8 GB 1067 MHz DDR3
memory. Figure~\ref{fig:poisson_cpu_time} gives the CPU time needed
for each case to converge in seconds, and
Figure~\ref{fig:poisson_iterations} gives the number of iterations
needed for each case to converge relative to the problem size.
\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{Adjoint_Direct_CPU_Time_2.pdf}
  \caption{\sl CPU Time (s) to converge vs. Problem Size ($N$ for an
    $N \times N$ square mesh). Both the adjoint and direct solvers are
    used with the five point and nine point stencils. A CPU time
    speedup is noted with the adjoint method due to the higher density
    of random walk events in regions with a large residual.}
  \label{fig:poisson_cpu_time}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{Adjoint_Direct_Iterations_2.pdf}
  \caption{\sl Iterations to converge vs. Problem Size ($N$ for an $N
    \times N$ square mesh). Both the adjoint and direct solvers are
    used with the five-point and nine-point stencils. }
  \label{fig:poisson_iterations}
\end{figure}

@@

As an additional comparison, the convergence behavior of MCSA can be
analyzed using both the adjoint and direct solvers to detect any
performance benefits. To assess the convergence properties of MCSA
using each solver and stencil, the infinity norm of the residual
computed in Eq.~(\ref{eq:MCSA-residual}) was collected at each
iteration for a fixed problem size of
$N=500$. Figure~\ref{fig:poisson_convergence} gives the results of
these computations. First, it is worthy to note on the semilog plot
that we are indeed achieving the expected exponential convergence from
MCSA with both Monte Carlo solvers. Second, we note that using the
adjoint method with the same number of stochastic histores per MCSA
iteration gives a faster rate of converge for the same reasons as
above.
\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{Adjoint_Direct_Convergence_2.pdf}
  \caption{\sl Infinity norm of the solution residual vs. iteration
    number for a problem of size $N=500$. Both the adjoint and direct
    solvers are used with the five point and nine point stencils. A
    higher rate of convergence is observed for MCSA using the adjoint
    Monte Carlo solver as compared to the direct method when both
    solvers compute the same number of random walks per iteration.}
  \label{fig:poisson_convergence}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{Sequential_CPU_Time.pdf}
  \caption{\sl CPU Time (s) to converge vs. Problem Size ($N$ for an
    $N \times N$ square mesh). Both the Sequential Monte Carlo and
    MCSA solvers are used with the five point stencils and the adjoint
    Monte Carlo solver. The number of random walks was twice the
    number of discrete states in the system in order to ensure
    convergence in the Sequential Monte Carlo method.}
  \label{fig:seq_cpu_time}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{Sequential_Iterations.pdf}
  \caption{\sl Iterations to converge vs. Problem Size ($N$ for an $N
    \times N$ square mesh). Both the Sequential Monte Carlo and MCSA
    solvers are used with the five point stencils and the adjoint
    Monte Carlo solver.}
  \label{fig:seq_iterations}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{Sequential_Convergence.pdf}
  \caption{\sl Infinity norm of the solution residual vs. iteration
    number for a problem of size $N=100$. Both the Sequential Monte
    Carlo and MCSA solvers are used with the five point stencils and
    the adjoint Monte Carlo solver.}
  \label{fig:seq_100}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{Sequential_Convergence.pdf}
  \caption{\sl Infinity norm of the solution residual vs. iteration
    number for a problem of size $N=500$. Both the Sequential Monte
    Carlo and MCSA solvers are used with the five point stencils and
    the adjoint Monte Carlo solver.}
  \label{fig:seq_500}
\end{figure}

>>>>

A timing and convergence study is used to demonstrate the
effectiveness of the adjoint method as compared to the direct
method. To assess both the CPU time and number of iterations required
to converge to a solution, a problem of constant $\Delta$ was used
with varying values of grid size, fixing the spectral radius of the
system at a constant value for each variation. Both the five-point and
nine-point stencils were used with both the direct and adjoint
solvers. For each case, $N \times N$ total random walk permutations
were computed per MCSA iteration where $N \times N$ is the number of
discrete grid points in the system. Solver parameters were set to a
weight cutoff of \sn{1}{-4} for the stochastic linear solver and a
convergence tolerance of \sn{1}{-8} for the MCSA iterative solver. All
computations presented in this section were completed on a 2.66 GHz
Intel Core i7 machine with 8 GB 1067 MHz DDR3
memory. Figure~\ref{fig:poisson_cpu_time} gives the CPU time needed
for each case to converge in seconds, and
Figure~\ref{fig:poisson_iterations} gives the number of iterations
needed for each case to converge relative to the problem size.
\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{dir_adj_cpu.pdf}
  \caption{\sl CPU Time (s) to converge vs. Problem Size ($N$ for an
    $N \times N$ square mesh). Both the adjoint and direct solvers are
    used with the five point and nine point stencils. A CPU time
    speedup is noted with the adjoint method due to the higher density
    of random walk events in regions with a large residual.}
  \label{fig:poisson_cpu_time}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{dir_adj_iterations.pdf}
  \caption{\sl Iterations to converge vs. Problem Size ($N$ for an $N
    \times N$ square mesh). Both the adjoint and direct solvers are
    used with the five-point and nine-point stencils. }
  \label{fig:poisson_iterations}
\end{figure}

@@

As an additional comparison, the convergence behavior of MCSA can be
analyzed using both the adjoint and direct solvers to detect any
performance benefits. To assess the convergence properties of MCSA
using each solver and stencil, the infinity norm of the residual
computed in Eq.~(\ref{eq:MCSA-residual}) was collected at each
iteration for a fixed problem size of
$N=500$. Figure~\ref{fig:poisson_convergence} gives the results of
these computations. First, it is worthy to note on the semilog plot
that we are indeed achieving the expected exponential convergence from
MCSA with both Monte Carlo solvers. Second, we note that using the
adjoint method with the same number of stochastic histores per MCSA
iteration gives a faster rate of converge for the same reasons as
above.
\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{dir_adj_convergence.pdf}
  \caption{\sl Infinity norm of the solution residual vs. iteration
    number for a problem of size $N=500$. Both the adjoint and direct
    solvers are used with the five point and nine point stencils. A
    higher rate of convergence is observed for MCSA using the adjoint
    Monte Carlo solver as compared to the direct method when both
    solvers compute the same number of random walks per iteration.}
  \label{fig:poisson_convergence}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{seq_cpu.pdf}
  \caption{\sl CPU Time (s) to converge vs. Problem Size ($N$ for an
    $N \times N$ square mesh). Both the Sequential Monte Carlo and
    MCSA solvers are used with the five point stencils and the adjoint
    Monte Carlo solver. The number of random walks was twice the
    number of discrete states in the system in order to ensure
    convergence in the Sequential Monte Carlo method.}
  \label{fig:seq_cpu_time}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{seq_iterations.pdf}
  \caption{\sl Iterations to converge vs. Problem Size ($N$ for an $N
    \times N$ square mesh). Both the Sequential Monte Carlo and MCSA
    solvers are used with the five point stencils and the adjoint
    Monte Carlo solver.}
  \label{fig:seq_iterations}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{seq_conv_100.pdf}
  \caption{\sl Infinity norm of the solution residual vs. iteration
    number for a problem of size $N=100$. Both the Sequential Monte
    Carlo and MCSA solvers are used with the five point stencils and
    the adjoint Monte Carlo solver.}
  \label{fig:seq_100}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=5in,clip]{seq_conv_500.pdf}
  \caption{\sl Infinity norm of the solution residual vs. iteration
    number for a problem of size $N=500$. Both the Sequential Monte
    Carlo and MCSA solvers are used with the five point stencils and
    the adjoint Monte Carlo solver.}
  \label{fig:seq_500}
\end{figure}
